---
title: "formative.rmd"
author: "Nick Anderson"
date: "October 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load the tidyverse library
library(tidyverse)
library(stringr)
```

4. Upload the MITx/Harvardx dataset to R as a dataframe named `D1`(Please inlcude your code in your markdown document)
```{r 4}
#load dataset
D1 <- read.csv("HMXPC13_DI_v2_5-14-14.csv", header = TRUE)
head(D1)
```


5. What does the `DI` suffix that is appeneded to two of the variable names mean?   
I am not sure, maybe, "Data Input"?
DI = de-identified, found on link

6. How many people enrolled in HarvardX/CS50x recieved certification in 2012? (Please inlcude your code in your markdown document)
```{r 6}
#filter data set to only view those who recieved certification in 2012
D2 <- filter(D1, certified == 1, str_detect(last_event_DI, "2012"))
head(D2[,c(6,13)])
sum(D2[,6]) #347 people recieved certification in 2012

#from the professor
D1b <- filter(D1, course_id == "HarvardX/CS50x/2012", certified == 1)
dim(D1b)
#correct answer is 1287
```

7. How many people registered for HarvardX/CS50x from Columbia on October 15, 2012? (Please inlcude your code in your markdown document)
```{r 7}
D3 <- filter(D1, str_detect(start_time_DI, "2012-10-15"), str_detect(course_id, "HarvardX/CS50x"), str_detect(final_cc_cname_DI, "Colombia"))
D3[,c(1, 7, 12 )] #10 students regiserted from Colombia for Harvardx/CS50x on October 15, 2012
```


8. Create a new data frame (`D2`) that includes student ID and registration date for all students from a country of your choice (Please inlcude your code in your markdown document)
```{r 8}
D2 <- D1 %>% filter(final_cc_cname_DI == "Portugal") %>%
  select(userid_DI, start_time_DI)
head(D2)
```


9. Create a new data frame (`D3`) in which the column names are the registration dates, with one column containing the student ids and all other cells indicating a `1` if the student registered on that date or an `NA` if they did not (hint: you will need to create a new variable before you change the structure of your data frame)
```{r 9}
mat <- matrix(data = NA, nrow = 476532, ncol = 413)
colnames(mat) <- unique(D1$start_time_DI)
users <- unique(D1$userid_DI)
DF3 <- cbind(users, mat)
DF3 <- as.data.frame(x)
DF3[1:3, 1:3]
#can't figure out... i was trying the spread function but cant get it to work

#Code
D3 <- D1[ , c("start_time_DI", "userid_DI", "registered")]
D3$registered <- as.character(D3$registered)
#register <- c(rep(NA, nrow(D3)))
#D3 <- data.frame(D3, register)
tag.D3 <- 0:1
names(tag.D3) <- c('NA', '1')
D3 <-  D3 %>% mutate(register.status = tag.D3[registered])
head(D3)

#from the professor, we shoujld use D2
D2$register <- 1
#Depending on the country you choose there are duplicate entries for registration and ID (someone should probably tell EdX about that :/)
D2 <- unique(D2)
D2 <- spread(D2, start_time_DI, register)
View(D2)
```

10. Now make the row names of `D3` the student ids on your new data frame and remove the user id column from your data frame
```{r 10}
row.names(D3) <- D3$userid_DI
D3$userid_DI <- NULL
View(D3)
```

11. Convert your data frame to a matrix (`M1`)
```{r 11}
M1[is.na(M1)] <- 0
M1 <- as.matrix(D3)
View(M1)
```


12. Create a new matrix which is the transposed version of your matrix (`M2`)
```{r 12}
M2 <- t(M1)
```

13. Create a third matrix (`M3`) that is a student by student matrix where the cells in the matrix represent if two studentes registered on the same day (This will take a lot of processing power, you may not be able to complete it if you have a large number of students)
```{r 13}
M3 <- M1 %*% t(M1)
```

14. Convert the diagonal of your matrix into `NA`s
diag(mat) <- NA
```{r}
diag(M3) <- NA
```